---
title: "RMarkdown Generation"
author: "Michael Mayer"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float:
      smooth_scroll: false
subtitle: Demo
---

```{r setup, warning = FALSE, message = FALSE, echo = FALSE}
library(tidyverse)
library(plotly)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE) 

#==================================================================
# GENERIC FUNCTIONS
#==================================================================

# Renderer (can be customized)
render <- function(x) {
  if(inherits(x, "character")) {
    cat(x)
  } else if (inherits(x, "ggplot")) {
    print(x)
  } else if (inherits(x, "htmlwidget")) {
    cat(htmltools::renderTags(as_widget(x))$html)
  } else if (inherits(x, "data.frame")) {
    print(knitr::kable(x, format = "pipe"))
  } else {
    stop("Cannot render object of class(es) ", class(x))
  }
  cat("\n\n")
}

# Solves problem with plotly and rmarkdown (call at the end)
# https://stackoverflow.com/questions/49990653/plotly-plot-doesnt-render-within-for-loop-of-rmarkdown-document
plotly_resolver <- function(some_list) {
  deps <- lapply(
    Filter(f = function(x) inherits(x, "htmlwidget"), x = some_list),
    FUN = function(hw) htmltools::renderTags(hw)$dependencies
  )
  htmltools::attachDependencies(
    x = htmltools::tagList(), 
    value = unlist(deps, recursive = FALSE)
  )
}

#==================================================================
# ANALYSIS (usually in a R-Script)
#==================================================================

# Perform the analysis and store results in nested list
data_list <- list(
  iris = list(data = iris, y = "Sepal.Length", x = "Species"),
  mtcars = list(data = mtcars, y = "mpg", x = "gear"),
  CO2 = list(data = CO2, y = "uptake", x = "Treatment")
)

results <- list()

for (nm in names(data_list)) { # nm <- "iris"
  X <- data_list[[nm]]
  results[[nm]]$n <- nrow(X$data)
  results[[nm]]$top6 <- head(X$data)
  p <- ggplot(X$data, aes_string(X$x, y = X$y)) +
    geom_jitter(color = "orange", alpha = 0.8, width = 0.1, height = 0)
  results[[nm]]$scatter <- p
  results[[nm]]$interactive <- ggplotly(p)
}

```

# Introduction

One of the most powerful reporting techniques in RMarkdown is to generate also the Markdown code (especially headers and navigation bar) from within R. It is the purpose of this notebook to demonstrate this.

The code does three things:

1. Keep analysis results in a nested list `results`. Usually, this part is done before heading for RMarkdown.
2. Move the contents of `results` to a (non-nested) list `out` along with additional markdown elements. The order of elements will be the one appearing in the resulting document.
3. Render each element in `out` with its corresponding render method.

# Results

```{r, results="asis", fig.keep="all", message=FALSE, warning=FALSE}

# Move content of "results" to flat list "out", along with headers
out <- list()

for (nm in names(results)) { # nm <- "iris"
  res <- results[[nm]]
  out[[length(out) + 1]] <- paste0("## Data `", nm, "`")  
  
  # Some text
  out[[length(out) + 1]] <- paste("The dataset has", res$n, "rows.")

  # A table with subheader
  out[[length(out) + 1]] <- paste("###", "Top six rows")  
  out[[length(out) + 1]] <- res$top6
  
  # A scatter plot with subheader
  out[[length(out) + 1]] <- paste("###", "A jittered scatter plot")  
  out[[length(out) + 1]] <- res$scatter
  
  # An interactive plot with subheader
  out[[length(out) + 1]] <- paste("###", "A corresponding interactive plot")  
  out[[length(out) + 1]] <- res$interactive
}

# Render each element
for (element in out) {
  render(element)
}

# plotly fix
plotly_resolver(out)
```
